\begin{abstract}
Large language model (LLM) agents that integrate external tools are vulnerable to indirect prompt injection (IPI), where malicious instructions embedded in tool-returned content hijack the agent. A growing body of work proposes defenses, yet these are often evaluated with proxy metrics---string matching for security, ``unblocked'' rates for utility---that can mislead conclusions. We systematize 10 IPI defenses plus a no-defense baseline across three classes (input encoding, structured query, runtime checking) and evaluate them on AgentDojo using \emph{native end-to-end evaluators}: the agent executes real tools, security is measured by \texttt{security\_from\_traces} on actual function-call traces, and utility by \texttt{user\_task.utility} on real environment state. Our evaluation spans benign (L0), static-attack (L1), and a capability-monotonic best-of-$k$ matched-suffix attack set (L2) that always contains the static payload, totaling 6{,}930 attack attempts plus benign and ablation trials. Three findings emerge. First, input-encoding defenses achieve the best security-utility tradeoff: their worst-case (best-of-$k$) Robust ASR is 0.19--0.30 with benign utility 0.80--0.90, far below the proxy-era value of 1.00 for the same defenses. The best-of-$k$ set increases ASR over the static payload significantly (paired McNemar $p<0.005$), but the increase comes from a \emph{generic} authority suffix, not the defense-matched suffix (which yields ASR $\leq 0.01$); we therefore do \emph{not} claim a signal-targeting adaptive attacker, and note that a length-matched control rules out a pure length effect. Second, separating benign utility (no-defense: 0.90) from attack-condition utility (0.19) reveals that 71\% of the no-defense utility loss is attack-induced, not a model capability limitation---correcting a prior misattribution; runtime-checking defenses that reach low ASR do so by over-blocking (benign utility as low as 0.03). Third, an ablation of our P2 multi-signal prototype shows the echo-canary provides independent security value (ASR 0.03 alone) but all signal combinations sacrifice utility entirely. Except for AgentDojo's official Spotlighting implementation, these figures characterize our reimplementations rather than the original authors' systems, evaluated on GPT-4o-mini and the AgentDojo workspace suite. We release our framework, native evaluation pipeline, and attack implementations as open-source artifacts.
\end{abstract}
