\appendix
\section{Statistical Analysis}
\label{app:stats}

\Cref{tab:stats} reports the L1 (static-payload) ASR and Utility with 95\% Wilson confidence intervals, computed from the native end-to-end results. These are \emph{marginal} proportions over 90 attempts per defense (10 user tasks $\times$ 3 injection tasks $\times$ 3 seeds).

\paragraph{Paired L1-vs-Robust analysis.} Because L1 and every L2 candidate reuse the same (user task, injection task, seed) units, the comparison is paired. We therefore do \emph{not} rely on marginal Wilson intervals for the paired difference $\Delta=\text{Robust ASR}-\mathrm{ASR}_{L1}$; instead we report an exact McNemar test on the discordant pairs and paired bootstrap confidence intervals for $\Delta$, clustering the bootstrap by user task to account for the hierarchical structure (the same task recurs across seeds, so the 90 attempts are not independent). Because the L2 candidate set contains the exact L1 payload, $\Delta\geq 0$ by construction, and the McNemar test measures whether the set adds \emph{additional} successes beyond L1. \Cref{tab:stats} reports the results (analysis code: \texttt{sok\_ipl/native/monotonic\_stats.py}). The best-of-$k$ set significantly increases ASR over L1 for the input-encoding defenses, StruQ, IPIGuard, P3, and no-defense (McNemar $p<0.02$, cluster-bootstrap CI excludes zero), and does not significantly increase it for FATH, Task Shield, P1, and P2 (which suppress ASR only by over-blocking; \Cref{app:overblock}).

\begin{table}[h]
\centering
\small
\caption{Paired L1-vs-Robust analysis (n=90 per defense, 3 seeds). ASR$_{L1}$ = static payload; Robust = capability-monotonic worst case over the L2 candidate set (includes L1). $\Delta$ = Robust $-$ ASR$_{L1}$ with 95\% paired-bootstrap CI; ``Cluster CI'' clusters by user task; $p$ is the exact McNemar two-sided $p$-value.}
\label{tab:stats}
\begin{tabular}{lccll c}
\toprule
Defense & ASR$_{L1}$ & Robust & $\Delta$ [95\% CI] & Cluster CI & McNemar $p$ \\
\midrule
Spotlighting & 0.09 & 0.19 & +0.10 [+0.04, +0.17] & [+0.02, +0.20] & 0.004 \\
Polymorphic & 0.10 & 0.29 & +0.19 [+0.11, +0.28] & [+0.08, +0.32] & $<$0.001 \\
Mixture-of-Enc. & 0.10 & 0.30 & +0.20 [+0.12, +0.29] & [+0.07, +0.36] & $<$0.001 \\
StruQ (prompt) & 0.63 & 0.89 & +0.26 [+0.17, +0.36] & [+0.14, +0.36] & $<$0.001 \\
FATH & 0.00 & 0.00 & +0.00 [+0.00, +0.00] & [+0.00, +0.00] & 1.000 \\
Task Shield & 0.06 & 0.07 & +0.01 [+0.00, +0.03] & [+0.00, +0.03] & 1.000 \\
IPIGuard & 0.40 & 0.52 & +0.12 [+0.06, +0.19] & [+0.06, +0.20] & 0.001 \\
P1: Echo Canary & 0.02 & 0.04 & +0.02 [+0.00, +0.06] & [+0.00, +0.07] & 0.500 \\
P2: Orthogonal & 0.00 & 0.00 & +0.00 [+0.00, +0.00] & [+0.00, +0.00] & 1.000 \\
P3: Task Invariant & 0.42 & 0.50 & +0.08 [+0.02, +0.13] & [+0.01, +0.16] & 0.016 \\
No Defense & 0.63 & 0.86 & +0.22 [+0.14, +0.31] & [+0.12, +0.33] & $<$0.001 \\
\bottomrule
\end{tabular}
\end{table}

\section{P2 Ablation Detail}
\label{app:p2_ablation}

\Cref{tab:p2_ablation_app} details the P2 ablation (90 trials per configuration, same tasks/attacks/seeds as the main experiment). The canary provides independent security value (ASR 0.03 alone vs.\ 0.74 baseline), but all configurations sacrifice Utility entirely.

\begin{table}[h]
\centering
\small
\caption{P2 ablation: canary vs.\ alignment contributions (n=90 each).}
\label{tab:p2_ablation_app}
\begin{tabular}{lcc}
\toprule
Configuration & ASR [95\% CI] & Utility [95\% CI] \\
\midrule
Canary only & 0.03 [0.01, 0.09] & 0.00 [0.00, 0.04] \\
Alignment only & 0.09 [0.05, 0.17] & 0.04 [0.02, 0.11] \\
Combined (P2) & 0.00 [0.00, 0.04] & 0.00 [0.00, 0.04] \\
\bottomrule
\end{tabular}
\end{table}

\section{Over-Blocking Diagnosis}
\label{app:overblock}

To distinguish genuine defense over-blocking from attack-induced failure, we compare Task Shield's benign (L0) and attack-condition (L1) utility with the no-defense baseline. Without attack, no-defense completes 90\% of tasks (benign Utility 0.90); under attack, this drops to 19\% (attack-condition Utility 0.19). This 71-percentage-point drop is attack-induced failure, not model capability limitation. Task Shield's benign over-blocking is $O_D = 0.90 - 0.43 = 0.47$: it blocks 47\% of benign tasks even without any attack. Its attack-condition utility (0.02) is a further 0.41 below its benign utility (0.43); this additional drop reflects the joint effect of the attack and the defense. The key insight is that the L0 baseline is essential: without it, one might attribute Task Shield's low attack-condition utility to the model, but the benign comparison shows the defense itself over-blocks nearly half of all benign tasks.

\section{Omitted Prior Experiments}
\label{app:omitted}

The prior version of this paper (proxy-based evaluation) included backbone comparisons (GPT-4o, Qwen2.5-7B), cross-benchmark validation (InjecAgent), adaptive-attack-strength sweeps, and SecAlign evaluation. These were conducted under the string-matching proxy methodology and produced results that \emph{contradict} our native end-to-end findings (e.g., proxy reported Spotlighting R-ASR=1.00; native shows ASR=0.21). We have therefore \emph{removed} these proxy-based results from this revision rather than report numbers we cannot support with end-to-end evidence. We plan to re-run backbone, cross-benchmark, and SecAlign evaluations using the native pipeline in future work.

\section{Native Evaluation Pipeline}
\label{app:native}

Our evaluation uses AgentDojo v1.2.2's native benchmark infrastructure. Each defense is implemented as a \texttt{BasePipelineElement} that integrates into AgentDojo's \texttt{AgentPipeline} and \texttt{ToolsExecutionLoop}. The agent executes real tools (e.g., \texttt{send\_email}, \texttt{get\_calendar\_events}), producing real function-call traces and environment state changes. Security is scored by \texttt{injection\_task.security\_from\_traces(traces, pre\_env, post\_env)} and utility by \texttt{user\_task.utility(model\_output, pre\_env, post\_env)}. This eliminates the string-matching proxy used in prior work. All code is in \texttt{sok\_ipl/native/pipeline\_builder.py} and \texttt{scripts/run\_bestofk.py}.

\paragraph{Capability-monotonic L2 candidate set.} For each unit $u=(\text{defense}, \text{seed}, \text{user task}, \text{injection task})$ we build a candidate set
\[
\mathcal{A}_{L2}(u)=\{a_{L1}, a_{\text{matched}}, a_{\text{generic}}, a_{\text{control}}, a_{\text{para},0},\dots,a_{\text{para},k-1}\},
\]
where $a_{L1}$ is the exact AgentDojo \texttt{important\_instructions} payload (verified byte-for-byte), each other candidate is $a_{L1}$ with an additive suffix inserted at a single fixed boundary after the goal (so the malicious goal is preserved verbatim), $a_{\text{generic}}$ is a target-independent authority suffix, and $a_{\text{control}}$ is a semantically-inert filler suffix matched in character length to $a_{\text{matched}}$ and derived deterministically by SHA-256 (isolating content effects from length). The per-unit outcome is $\mathrm{success}(u)=\bigvee_{a\in\mathcal{A}_{L2}(u)} \mathrm{security\_from\_traces}(a)$, and
\[
\text{Robust ASR}=\operatorname*{mean}_u \max_{a\in\mathcal{A}_{L2}(u)} \mathrm{ASR}(a) \;\geq\; \mathrm{ASR}_{L1},
\]
since $a_{L1}\in\mathcal{A}_{L2}(u)$. The FATH matched candidate embeds the real $\mathrm{SHA\text{-}256}(\text{user task})[{:}8]$ token; the P2 matched candidate attacks both the canary and the alignment check. The runner (\texttt{scripts/run\_bestofk.py}) writes one attempt record per candidate and \texttt{sok\_ipl/native/monotonic\_stats.py} computes best-of-$k$, McNemar, and cluster bootstrap, asserting $\mathrm{l1\_success}\Rightarrow\mathrm{l2\_set\_success}$ for every unit.
