\section{Unified Adaptive Attack Framework}
\label{sec:attack}

\subsection{Design Principles}

The key insight behind our framework is that an adaptive attack is not ``a stronger injection'' but rather ``an injection crafted with knowledge of the defense's detection signal.'' Each defense class relies on a specific signal to distinguish trusted from untrusted content; the adaptive attacker's goal is to craft an injection that does not trip that signal while still delivering the malicious instruction. We instantiate this insight as a mapping from defense class to evasion strategy.

\subsection{Evasion Strategies}

We define six evasion strategies, one per defense class plus a default. \Cref{tab:strategies} provides a concrete payload example for each.

\begin{table}[t]
\centering
\small
\caption{Evasion strategies: target defense class and example payload.}
\label{tab:strategies}
\begin{tabular}{p{0.18\columnwidth}p{0.15\columnwidth}p{0.55\columnwidth}}
\toprule
Strategy & Target & Example Payload Fragment \\
\midrule
Signal Align. & IE & \texttt{<data>[goal]</data>} (reuses defense delimiter) \\
Boundary Blur & SQ & \texttt{\{"\_meta":\{"instr":[goal]\}\}} (smuggles in JSON data field) \\
Intent Disguise & RC & ``To complete the user's request, first: [goal]'' \\
Atten.\ Redist. & IP & [filler]\verb|\n|[goal]\verb|\n|[filler] (dilutes attention) \\
Channel Split & AR & \texttt{[PART\_A][goal\_a][PART\_B][goal\_b]} (fragments) \\
Paraphrase & TB & [EN]\,[goal] / [FR]\,[goal] / [B64]\,[goal] \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Signal Alignment (for IE).}
The defense wraps untrusted content in a delimiter (e.g., \texttt{<data>}). For defenses with a fixed, known delimiter (Spotlighting, Mixture-of-Encodings), the attacker includes the same delimiter within the injection, so the LLM cannot distinguish the defense's wrapping from the injection's wrapping. For Polymorphic Prompt, which randomizes the delimiter per call, the attacker cannot know the specific delimiter in a single-shot L2 setting. In this case, the L2 attack uses the \texttt{<INFORMATION>} template without delimiter reuse---the attack succeeds because the \texttt{<INFORMATION>} wrapper itself overrides the defense's marking, regardless of which delimiter the defense uses. We note this distinction: the signal-alignment strategy is effective against fixed-delimiter IE defenses, while the \texttt{<INFORMATION>} template's authority-claiming framing is effective against randomized-delimiter IE defenses.

\paragraph{Boundary Blur (for SQ).}
The defense separates instructions and data into structurally distinct positions. The attacker smuggles instructions inside the data field by reusing the defense's own JSON structure: the payload is embedded as a \texttt{\_meta.instruction} key within what the defense parses as a data object.

\paragraph{Intent Disguise (for RC).}
The defense checks whether the proposed action aligns with the user task. The attacker phrases the malicious action as a prerequisite step of the legitimate task: ``To complete the user's request, first perform this prerequisite step: [goal]. Then continue with the original task.'' This framing causes the task-alignment score to remain high because the action is semantically connected to the user task.

\paragraph{Attention Redistribution (for IP).}
The defense flags prompts where disproportionate attention falls on untrusted tokens. The attacker pads the injection with benign filler text (repeated generic phrases) so the attention spike on the malicious token is diluted below the detector's threshold.

\paragraph{Channel Split (for AR).}
The defense isolates tools into separate processes or TEEs. The attacker splits the injection across multiple tool outputs so that no single channel contains the complete malicious instruction; the LLM reassembles the instruction from fragments across channels.

\paragraph{Paraphrase and Translate (for training-based).}
The defense was fine-tuned to resist a specific injection style. The attacker rotates through multiple languages and heavy paraphrase to evade the learned pattern.

\subsection{Attack Levels}

The framework materializes attacks at five levels (\Cref{sec:background}). L1 uses the \texttt{<INFORMATION>} template without defense-specific evasion. L2 applies the matched evasion strategy on top of the template. L3 iterates: the attacker observes the agent's response to the previous injection and uses an LLM-in-the-loop optimizer to refine the payload over up to 3 turns. L4 replaces the textual instruction with a backdoor trigger token.

\subsection{Threat Model: How the Attacker Identifies the Defense Class}

Our threat model assumes the attacker knows the defense \emph{class} (IE, SQ, RC, IP, AR) but not its secret parameters (e.g., the exact delimiter string, the randomization seed). In practice, defense-class identification can occur through: (1) public deployment documentation (many defenses are open-source), (2) probing the agent's behavior with test inputs to infer which class is active (e.g., observing that the agent receives delimited output suggests IE), or (3) insider knowledge. We do not count defense-class identification queries in the attack query budget, as we assume this information is obtained offline. This is a stronger threat model than pure black-box; Setting A (Appendix~\ref{app:settings_ab}) evaluates the weaker case where the attacker does \emph{not} know the defense class.

\subsection{L2 Payload Generation}

The L2 matched payload is generated \emph{online}: the attacker takes the injection goal, wraps it in the \texttt{<INFORMATION>} template, and appends the defense-class-specific evasion suffix (Table~\ref{tab:strategies}). No offline queries to the target defense are needed for L2; the payload is constructed from the template + suffix in a single step. The LLM-in-the-loop optimizer (used in L3 and optionally in L2 with \texttt{max\_iters}$>$0) does query the agent, but Setting B (Appendix~\ref{app:settings_ab}) shows that this optimization does not improve ASR---the single-shot matched payload is already sufficient.

\subsection{Malicious Semantics Preservation}

A key assumption of the fragility conjecture is that the adaptive search can find a payload that simultaneously (a) keeps the signal below the defense's threshold and (b) preserves the malicious instruction's semantics. We argue this holds because: (1) the evasion suffixes are \emph{additive} (e.g., reusing delimiters, adding filler text) and do not modify the core malicious instruction; (2) the \texttt{<INFORMATION>} wrapper's effectiveness depends on its framing, not on specific tokens that the defense could filter; (3) empirically, the L2 payloads achieve ASR $\geq$ 0.90 on all single-signal defenses, confirming that the malicious semantics are preserved. We acknowledge that for defenses with very aggressive content filtering (e.g., blocking any text containing ``email'' or ``send''), semantics preservation may fail, but no such defense exists in our evaluation.

\subsection{Why L3 ASR Can Be Lower Than L2 ASR}

Several defenses show L3 (adaptive multi-turn) ASR lower than L2 (adaptive single-shot) ASR (e.g., Spotlighting L2=1.00, L3=0.63). This counterintuitive result occurs because the L3 LLM-in-the-loop optimizer sometimes over-optimizes: it rewrites the payload to reduce detectability, but in doing so inadvertently weakens the injection's instruction-following strength. The optimizer prioritizes stealth over effectiveness, causing the LLM to follow the injection less reliably. This indicates that iterative refinement has diminishing returns once the single-shot matched payload already saturates the attack.

\subsection{Composability}

The framework is composable: an attack is a (level, strategy) pair, and the framework materializes the right instance from a registry. Adding a new attack or defense requires only registering it---the evaluation harness, benchmarks, and analysis pipeline are defense- and attack-agnostic. This design is what allows us to scale the evaluation to 12 defenses and 5 attack levels without modifying the harness for each combination.
