\section{Conclusion}
\label{sec:conclusion}

We presented a systematization of knowledge on indirect prompt injection (IPI) defenses for LLM agents, evaluated end-to-end. Using AgentDojo's native evaluators---measuring real tool-execution traces for security and real environment state for utility---we evaluated 10 defenses plus a no-defense baseline across three classes in 2{,}580 trials spanning benign (L0), static-attack (L1), and matched-suffix (L2) conditions. Our central contribution is that \emph{end-to-end evaluation changes the ranking of IPI defenses}: input-encoding defenses achieve the best security-utility tradeoff, with a capability-monotonic worst-case (best-of-$k$) Robust ASR of 0.19--0.30 at benign utility 0.80--0.90, whereas a string-matching proxy reported ASR 1.00 for the same defenses on the same setting. Our Robust ASR is computed over a matched-suffix candidate set that always includes the static payload, so it is $\geq$ the static ASR (0.09--0.10) by construction; the best-of-$k$ increase is statistically significant (paired McNemar $p<0.005$) but comes from a generic authority suffix rather than the defense-matched suffix (ASR $\leq 0.01$). We explicitly do \emph{not} claim to have built a signal-targeting adaptive attacker or that these defenses resist one---establishing genuine adaptive robustness would require optimization- or search-based attackers beyond our fixed candidate set. Second, by measuring benign utility separately (no-defense: 0.90), we show that 71\% of the attack-condition utility loss is attack-induced failure, not a model capability limitation; defining benign over-blocking as $O_D=U_{\text{no-def},L0}-U_{D,L0}$, input-encoding defenses over-block 0--0.10 while runtime-checking defenses over-block 0.47--0.90. Third, a P2 ablation shows the echo-canary provides independent security value (ASR 0.03 alone) but all signal combinations sacrifice utility entirely. These results, obtained on GPT-4o-mini and the AgentDojo workspace suite, demonstrate that end-to-end evaluation with native evaluators and benign baselines is essential for honestly characterizing IPI defense tradeoffs; broader models, suites, and search-based attackers are important future work. We release our framework, native evaluation pipeline, and attack implementations as artifacts.
