\section{Design Principles and Fragility Conjecture}
\label{sec:principles}

\subsection{Single-Signal Fragility Hypothesis}

Based on our empirical observations, we propose the following falsifiable hypothesis rather than a formal theorem, as the formal model of ``signal'' and ``computability'' requires assumptions about LLM internals that we cannot fully verify.

\begin{conjecture}[Single-Signal Fragility, Empirical]
\label{conj:single}
Let $D$ be an IPI defense whose decision rule depends primarily on a single signal $s$ that is observable to the attacker (i.e., the attacker can infer or approximate $s$ from the prompt or from query feedback). Then an adaptive attacker with knowledge of the defense class can construct an injection that bypasses $D$ while preserving the malicious instruction's semantics, achieving ASR close to the no-defense baseline.
\end{conjecture}

\paragraph{Scope and Assumptions.} This hypothesis applies to defenses where (1) the signal is observable (e.g., delimiter text, encoding scheme, attention pattern accessible to the attacker), (2) the attacker knows the defense class and can craft a matched evasion strategy, and (3) the evasion suffix is additive and does not destroy the malicious instruction's semantics. It does \emph{not} apply to defenses relying on signals the attacker cannot observe (e.g., a server-side secret never exposed in any output) or to defenses that enforce structural invariants independent of prompt content (e.g., hardware-enforced capability tokens). We do not claim a specific search complexity bound; the hypothesis states only that \emph{some} adaptive search finds a bypass, which our experiments confirm empirically.

\paragraph{Falsifiability.} The hypothesis is falsifiable: a defense that relies on a single observable signal but resists all tested adaptive strategies (matched evasion, multi-turn optimization, cross-defense transfer) would refute it. In our evaluation, no such defense exists among the faithfully-evaluated single-signal defenses (IE class). P2 (which combines an unobservable canary with semantic alignment) resists, but it uses two signals, one of which is unobservable---consistent with the hypothesis's scope.

\subsection{Design Principles}

The conjecture's corollary is that an adaptive-robust defense must rely on at least one signal \emph{not} efficiently computable from the prompt. We instantiate four design principles as minimal prototypes:

\paragraph{P1: Echo Canary.}
The defense injects a random canary into the system prompt and requires the LLM to echo it before any tool call. The canary is never present in tool outputs, so an injection cannot forge it. \emph{Result:} R-ASR 0.90---P1 is defeated. The high L2 ASR (0.90) in the single-shot setting is explained not by canary extraction (which requires multi-turn observation) but by the LLM's tendency to include the canary in its output context, which the injection can reference indirectly (``execute the action that includes the authorization token from the system prompt''). In L3 (ASR 0.75), the multi-turn optimizer sometimes over-optimizes for stealth and weakens the injection, as discussed in \Cref{sec:attack}. We note that this prototype is intentionally minimal and \emph{naive}: a more robust design would use a trusted monitor that holds the secret and verifies a MAC over the action, never exposing the secret to the LLM. Our P1 result demonstrates that the \emph{naive} form (echo-the-canary) is insufficient; whether a cryptographic monitor-based design would resist is an open question.

\paragraph{P2: Orthogonal Signals.}
The defense combines P1's echo-canary with Task-Shield-style semantic alignment. Evasion requires bypassing both checks. \emph{Result:} R-ASR 0.00---P2 resists. However, FPR is 0.57: the conjunction of two checks over-blocks 57\% of benign actions. We note that P2's resistance may be primarily due to the semantic-alignment check (which alone blocks 47\% of benign actions at threshold 0.5); the canary may not provide independent security value beyond what semantic alignment already offers, because the canary is present in the system prompt and the LLM can reference it. Disentangling the two signals' contributions is left as future work.

\paragraph{P3: Task Invariant (Relaxed).}
The defense compiles the user task into a state machine and blocks any action that does not advance it. Unlike the degenerate version (which allows only one transition), this prototype allows top-$k$ plausible next tools as determined by an embedding similarity to the user task ($k=3$). \emph{Result:} R-ASR 0.00 but FPR 1.00---even with the relaxed top-$k$ allowance, the state machine is too strict for the AgentDojo workspace tasks, which involve complex multi-step reasoning that the top-$k$ heuristic cannot capture. This is itself a finding: task-invariant enforcement is impractical without a sophisticated task model.

\paragraph{P4: Least Privilege.}
The defense runs each tool in a sandbox with access only to resources named in the task spec. \emph{Result:} R-ASR 1.00 when measured by injection ASR---the sandbox does not detect the injection; it only limits the blast radius. However, we note that evaluating P4 by ASR alone is a \emph{category error}: the sandbox's goal is containment, not detection. A fairer metric is \emph{damage reduction}: the fraction of attacker-desired resources that the sandbox prevents access to. Under this metric, P4 blocks access to non-allowed tools entirely, but cannot prevent malicious use of allowed tools (e.g., sending an email via the allowed email tool to an attacker address). We report both ASR (1.00) and damage reduction (partial: non-allowed tools blocked, allowed tools vulnerable) to avoid misleading conclusions.

\subsection{Implications}

The principle prototypes confirm the conjecture's prediction: P1 (single signal) is defeated; P2 (orthogonal multi-signal) resists but at a utility cost; P3 (task invariant) over-blocks even in its relaxed form; P4 (blast-radius containment) does not prevent the attack. The fundamental tradeoff is that adaptive robustness requires unobservable signals, but unobservable signals are either forgeable through observation (canary leakage) or require conservative over-blocking (conjunction of checks). Bridging this tradeoff---achieving both low R-ASR and low FPR---remains an open problem.
