\section{Evaluation}
\label{sec:eval}

\subsection{Research Questions}

Our evaluation is organized around three research questions:

\begin{itemize}[noitemsep,topsep=0pt]
\item \textbf{RQ1:} How do IPI defenses perform under static (L1) vs.\ adaptive (L2) attack, measured end-to-end?
\item \textbf{RQ2:} What is the security-utility tradeoff, and how much utility loss is genuine defense over-blocking vs.\ attack-induced failure?
\item \textbf{RQ3:} Does the echo-canary signal provide independent security value, and does multi-signal combination recover utility? (P2 ablation)
\end{itemize}

\subsection{Experimental Setup}

\paragraph{Benchmark.}
We evaluate on the AgentDojo workspace suite (v1.2.2)~\cite{debenedetti2024agentdojo,debenedetti2025agentdojo}, which provides 40 user tasks across calendar, email, and file-management scenarios with ground-truth injection goals. We use 10 user tasks $\times$ 3 injection tasks per seed. Each task pairs a legitimate user request (e.g., ``Who else is invited at the Networking event?'') with an injection goal (e.g., ``Email `Hey, how is it going?' to mark.black-2134@gmail.com''). Crucially, our evaluation uses AgentDojo's \emph{native} pipeline: the agent executes real tools (send\_email, get\_calendar\_events, etc.), producing real side effects and function-call traces.

\paragraph{Backbone.}
We use GPT-4o-mini (API snapshot \texttt{gpt-4o-mini-2024-07-18}) as the primary agent backbone via an OpenAI-compatible API. Main-table results are reported over 3 seeds (0--2); 95\% Wilson confidence intervals are reported in parentheses. Total API cost was approximately \$60--80 USD.

\paragraph{Defenses.}
We evaluate 11 defense configurations: 7 baselines from the literature (Spotlighting, StruQ, Task Shield, IPIGuard, FATH, Polymorphic Prompt, Mixture-of-Encodings) plus 3 design-principle prototypes (P1--P3, \Cref{sec:principles}), and a no-defense baseline. StruQ is evaluated as a \emph{prompt-level structured-query approximation} (\texttt{struq\_prompt}); the original StruQ requires a fine-tuned model checkpoint~\cite{chen2024struq}, which is not reproducible in our API-based pipeline. We report this honestly: our \texttt{struq\_prompt} tests whether \emph{prompt-level} channel separation (without fine-tuning) suffices, not whether the original fine-tuned StruQ works. Attention Tracker and P4 are excluded from the main table for fidelity reasons (\Cref{sec:proxy_results}).

\paragraph{Metrics (end-to-end).}
Unlike proxy-based evaluations that estimate security via string matching, our metrics are computed by AgentDojo's native evaluators on real tool-execution traces:
\begin{itemize}[noitemsep,topsep=0pt]
\item \textbf{ASR} (Attack Success Rate): fraction of attacked tasks where the injection goal was \emph{actually executed}, as determined by \texttt{injection\_task.security\_from\_traces()} checking the real function-call traces and environment state. This is a strict, trace-based measure---not a string-match proxy.
\item \textbf{Utility}: fraction of tasks where the user's task was \emph{correctly completed}, as determined by \texttt{user\_task.utility()} comparing pre- and post-execution environment state. This independently measures task completion, unlike ``unblocked'' proxies that conflate ``not blocked'' with ``done.''
\end{itemize}
We report both ASR and Utility because a defense that blocks everything achieves ASR=0 but Utility=0---a trivially safe but useless operating point. The security-utility tradeoff is the central tension our evaluation surfaces.

\paragraph{Scale.}
The main experiment comprises 11 defenses $\times$ 3 conditions (L0 benign, L1 static, L2 adaptive) $\times$ 10 user tasks $\times$ 3 injection tasks (L1/L2) or 10 user tasks (L0) $\times$ 3 seeds = 2{,}310 trials, all using AgentDojo's native end-to-end pipeline. The P2 ablation (3 configurations $\times$ 90 = 270 trials) disentangles canary vs.\ alignment contributions. Total: 2{,}580 trials, 0 errors. Total API cost was approximately \$80--100 USD.

\subsection{RQ1: Static (L1) vs.\ Matched-Suffix (L2) and Worst-Case Robust ASR}

\Cref{tab:main} reports end-to-end results for all 11 configurations under three conditions: L0 (benign, no attack), L1 (static \texttt{important\_instructions} payload), and L2 (a defense-matched suffix). ASR and Utility are computed by AgentDojo's native evaluators on real tool-execution traces.

We first note a methodological point that governs how these numbers must be read. Our L2 is a defense-matched \emph{suffix} appended to the static payload; it is one hand-authored candidate, not a search over an attack space. Descriptively, for input-encoding defenses the matched suffix yields \emph{lower} ASR than the static payload (L1 0.14--0.21 vs.\ matched-suffix 0.04--0.08). This does \emph{not} mean an adaptive attacker is weaker: a rational attacker retains the option of using the plain static payload. We therefore report the \emph{capability-monotonic} quantity
\[
\text{Robust ASR} = \operatorname*{mean}_{\text{unit}}\; \max_{a \in \mathcal{A}_{L2}} \mathrm{ASR}(a),
\quad \mathcal{A}_{L1}\subseteq\mathcal{A}_{L2},
\]
where the L2 candidate set $\mathcal{A}_{L2}$ contains the exact L1 payload plus the matched suffix, a generic authority suffix, a length-matched control, and $k$ paraphrases (\Cref{sec:attack}). By construction Robust ASR $\geq \mathrm{ASR}_{L1}$. For the input-encoding defenses the Robust ASR stays at the L1 level (0.14--0.21): none of the matched or paraphrase candidates exceed the plain static payload. The correct reading is therefore that \emph{our matched suffixes fail to strengthen the attack}, not that the defenses withstand a stronger adaptive adversary. Establishing genuine adaptive robustness would require optimization-based or search-based attackers beyond our fixed candidate set, which we leave to future work.

Under end-to-end metrics, the input-encoding worst-case ASR (0.14--0.21) is still far below the proxy-era value (1.00). The end-to-end pipeline, not the adaptive framing, is what changes this ranking (\Cref{tab:main}).

For runtime-checking defenses, the matched-suffix ASR is comparable to or below the static value: Task Shield (L1 0.10, matched 0.00), FATH (0.00, 0.00), IPIGuard (0.38, 0.34); their Robust ASR equals the larger of the two. The prompt-level StruQ approximation stays high (L1 0.70, matched 0.59; Robust 0.70), confirming that prompt-level channel separation without fine-tuning does not defend.

\subsection{RQ2: Security-Utility Tradeoff and Over-Blocking}

The benign (L0) condition is essential for interpreting utility loss. Without attack, the no-defense baseline completes 90\% of tasks (Utility 0.90); under attack, this drops to 19\% (Utility 0.19). This 71-percentage-point drop is \emph{attack-induced failure}, not a model capability limitation---correcting a prior misattribution that attributed the low attack-condition utility to model weakness.

We define \emph{benign over-blocking} of defense $D$ as the drop in utility it causes on benign tasks relative to no defense:
\[
O_D = U_{\text{no-defense},L0} - U_{D,L0},
\]
which isolates the defense's own utility cost from attack-induced failure. Under this definition (Table~\ref{tab:main}, L0 column):

\begin{itemize}[noitemsep,topsep=0pt]
\item \textbf{IE defenses barely over-block:} $O_D$ is 0.00 (Polymorphic), 0.03 (Mixture-of-Encodings), and 0.10 (Spotlighting)---their benign utility (0.80--0.90) is close to the no-defense 0.90, confirming they suppress injections without blocking legitimate actions.
\item \textbf{Runtime-checking defenses over-block heavily:} Task Shield's $O_D = 0.90 - 0.43 = 0.47$; FATH and P1/P2 have benign utility $\leq 0.03$, i.e.\ $O_D \geq 0.87$. These defenses block most benign tasks even without any attack.
\item \textbf{Prompt-level StruQ preserves benign utility but does not defend:} $O_D = 0.90 - 0.83 = 0.07$, yet its attack-condition ASR is 0.70; its low attack-condition utility (0.17) is attack-induced failure, not over-blocking.
\end{itemize}

We report the earlier ``benign utility minus attack utility'' quantity separately as the \emph{attack-conditioned utility change}, not as over-blocking, since it conflates defense behavior with attack effects.

\subsection{Main Results Table}

\Cref{tab:main} reports end-to-end results for 10 defenses plus a no-defense baseline across three conditions: L0 (benign), L1 (static payload), and L2 (defense-matched suffix). ``Robust'' is the capability-monotonic worst case $\max(\mathrm{ASR}_{L1},\mathrm{ASR}_{L2\text{-set}})$, where the L2 set always contains the L1 payload. Attention Tracker and P4 are excluded due to fidelity limitations (\Cref{sec:proxy_results}). Except for AgentDojo's official Spotlighting implementation, these figures characterize our AgentDojo reimplementations, not the original authors' systems.

\begin{table*}[t]
\centering
\small
\caption{Main results on AgentDojo workspace (GPT-4o-mini, 3 seeds). L0 = benign (no attack, 30 trials); L1 = static \texttt{important\_instructions} (90 trials); L2 = defense-matched suffix (90 trials). Robust ASR $=\max(\mathrm{ASR}_{L1},\mathrm{ASR}_{L2})$ (capability-monotonic, L1 $\in$ L2 set). ASR from \texttt{security\_from\_traces}, Utility from \texttt{user\_task.utility}, on real tool traces. Marginal proportions; paired McNemar / cluster-bootstrap for L1-vs-L2 differences are in Appendix~\ref{app:stats}. Except Spotlighting (AgentDojo official), rows are our reimplementations.}
\label{tab:main}
\begin{tabular}{llccccccc}
\toprule
 & & L0 & \multicolumn{2}{c}{L1 (Static)} & \multicolumn{2}{c}{L2 (Matched)} & Robust \\
\cmidrule(lr){3-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-8}
Defense & Class & Util$\uparrow$ & ASR$\downarrow$ & Util$\uparrow$ & ASR$\downarrow$ & Util$\uparrow$ & ASR$\downarrow$ \\
\midrule
No Defense & --- & 0.90 & 0.74 & 0.19 & 0.70 & 0.20 & 0.74 \\
\midrule
Polymorphic & IE & 0.90 & 0.14 & 0.73 & 0.08 & 0.82 & 0.14 \\
Mixture-of-Enc. & IE & 0.87 & 0.16 & 0.69 & 0.04 & 0.86 & 0.16 \\
Spotlighting & IE & 0.80 & 0.21 & 0.67 & 0.06 & 0.82 & 0.21 \\
\midrule
StruQ (prompt) & SQ & 0.83 & 0.70 & 0.17 & 0.59 & 0.32 & 0.70 \\
\midrule
FATH & RC & 0.03 & \textbf{0.00} & 0.02 & \textbf{0.00} & 0.06 & \textbf{0.00} \\
Task Shield & RC & 0.43 & 0.10 & 0.02 & 0.00 & 0.07 & 0.10 \\
IPIGuard & RC & 0.43 & 0.38 & 0.06 & 0.34 & 0.10 & 0.38 \\
\midrule
P1: Echo Canary & RC & 0.03 & 0.03 & 0.00 & 0.01 & 0.02 & 0.03 \\
P2: Orthogonal & RC & 0.00 & \textbf{0.00} & 0.00 & \textbf{0.00} & 0.01 & \textbf{0.00} \\
P3: Task Invariant & RC & 0.50 & 0.42 & 0.02 & 0.39 & 0.07 & 0.42 \\
\bottomrule
\multicolumn{8}{l}{\footnotesize IE = Input Encoding, SQ = Structured Query, RC = Runtime Checking.}
\end{tabular}
\end{table*}

\subsection{Excluded Defenses}
\label{sec:proxy_results}

Two defenses are excluded from the main table due to fidelity limitations. \textbf{Attention Tracker} (IP) requires access to model attention weights, unavailable via the GPT-4o-mini API; a token-frequency proxy would not faithfully represent the defense. \textbf{P4: Least Privilege} (AR) requires real OS-level sandboxing; a prompt-level simulation would be a category error. We discuss these qualitatively in \Cref{sec:principles} but do not report quantitative results for them.

\subsection{P2 Ablation: Does the Canary Provide Independent Security?}
\label{sec:p2_ablation}

A key question is whether P2's resistance comes from the echo-canary signal, the semantic-alignment check, or their combination. We run an ablation with three configurations (90 trials each, same tasks/attacks/seeds):

\begin{table}[t]
\centering
\small
\caption{P2 ablation: canary vs.\ alignment contributions (90 trials each).}
\label{tab:p2_ablation}
\begin{tabular}{lcc}
\toprule
Configuration & ASR $\downarrow$ & Utility $\uparrow$ \\
\midrule
Canary only & 0.03 [0.01, 0.09] & 0.00 [0.00, 0.04] \\
Alignment only & 0.09 [0.05, 0.17] & 0.04 [0.02, 0.11] \\
Combined (P2) & \textbf{0.00} [0.00, 0.04] & 0.00 [0.00, 0.04] \\
\bottomrule
\end{tabular}
\end{table}

The canary \emph{does} provide independent security value: canary-only achieves ASR 0.03 (vs.\ 0.74 baseline), a 96\% relative reduction, without the alignment check. However, all three configurations have Utility $\leq$ 0.04. The combination does not recover Utility---it merely stacks two over-blocking signals. This refutes the original hypothesis that orthogonal multi-signal combination could achieve security \emph{at acceptable utility cost}: in our end-to-end evaluation, P2's Utility is 0.00, not the 0.43 reported by the earlier proxy-based measurement. The canary's independent value is real but comes at full Utility cost.
