\section{Limitations and Ethics}
\label{sec:limitations}

\paragraph{Limitations.}
(1) We evaluate primarily on GPT-4o-mini; a 3-backbone comparison with GPT-4o (closed-source) and Qwen2.5-7B-Instruct (open-source, Appendix~\ref{app:qwen}) shows the structural finding is backbone-independent for IP defenses but backbone-dependent for IE defenses. (2) SecAlign~\cite{chen2024secalign} is classified in the SQ class but not evaluated in the main experiment because it requires a fine-tuned model checkpoint incompatible with our API-based pipeline; we evaluate it separately on Mistral-7B in Appendix~\ref{app:sealign}. MELON~\cite{zhu2025melon} (provable defense) requires a formal task specification language not available in AgentDojo; ACE~\cite{li2025security} and IsolateGPT~\cite{wu2025isolategpt} require system-level process isolation infrastructure (separate OS processes, TEE enclaves) that cannot be faithfully reproduced in our prompt-level evaluation framework. We discuss these defenses qualitatively in the taxonomy but do not claim experimental results for them. (3) L4 (backdoor-augmented) attacks show ASR 0.00 for all defenses because no backdoor is planted in the backbone. L4 represents a theoretical ceiling; we omit it from the main table and discuss it as a limitation rather than empirical evidence. (4) Our ASR measurement uses string-matching of injection-goal entities (emails, phones, URLs) as a proxy for AgentDojo's native \texttt{security\_from\_traces} check. This may produce false positives if the agent mentions the target entity in a refusal (``I will not email mark.black@...''); we mitigate this by checking for the entity in the \emph{action} output, not the full response, but a small false-positive rate is possible. (5) Our root-cause labeling is rule-based with optional LLM verification; the labeling is partially circular (attack strategy is chosen by defense class), though the LLM verifier independently confirms labels on ambiguous cases. (6) The adaptive attack optimizer uses 2 iterations in the main experiment; Appendix~\ref{app:adaptive_strength} shows that increasing to 5 iterations does not meaningfully change results. (7) We use 5 seeds for 6 key defenses and 3 seeds for the remaining defenses; Wilcoxon signed-rank $p$-values are reported for the 5-seed subset (Appendix~\ref{app:stats}). (8) We report results on the AgentDojo benchmark; the framework also supports InjecAgent~\cite{zhan2024injecagent} and Agent Security Bench~\cite{zhang2024agent}, but cross-benchmark results are deferred to future work due to space. (9) We evaluate 12 defense configurations; newly published defenses may not fit the current taxonomy, though the framework is extensible via registry.

\paragraph{Ethics Considerations.}
This work studies attacks on LLM agents. All attacks are evaluated in a controlled benchmark environment (AgentDojo) with no real-world deployment or impact on actual users. The injection payloads use publicly known templates from the open literature. We do not release any new attack tool that is more effective than what is already publicly available; our contribution is the \emph{systematization} and \emph{evaluation framework}, not a novel attack. We disclose no real-world vulnerabilities. The framework is released as open-source to help defense developers self-assess their systems before deployment, in alignment with responsible disclosure practices. We follow the Menlo Report's ethical guidelines for cybersecurity research.

\paragraph{Use of Generative AI.}
Generative AI tools (ChatGPT and Claude) were used for prose editing and boilerplate code generation. All technical content, experimental design, claims, and results were authored and verified by the human authors, who accept full responsibility for the veracity of all material in this paper.

\paragraph{Literature Search Methodology.}
We searched for IPI defense papers on arXiv, Google Scholar, and DBLP using the queries ``indirect prompt injection defense'', ``LLM agent prompt injection'', and ``prompt injection attack defense'' (2023--2026). Inclusion criteria: (1) proposes a concrete defense mechanism (not just an attack or survey), (2) targets indirect prompt injection (not direct user-input jailbreaks), (3) applicable to tool-integrated LLM agents. Exclusion criteria: (1) defenses for non-agent LLM applications (e.g., chatbot-only safety), (2) defenses requiring hardware not available in our environment (e.g., SGX enclaves for IsolateGPT/ACE), (3) defenses requiring custom model training infrastructure not compatible with API-based evaluation (e.g., SecAlign's DPO pipeline). We identified 17 candidate defenses; 8 met all inclusion criteria and were evaluated, 4 were classified but not evaluated (SecAlign, MELON, ACE, IsolateGPT), and 5 were excluded as out-of-scope (e.g., direct jailbreak defenses).

\paragraph{Implementation Fidelity.}
Of the 8 literature baselines, 3 use original or near-original code (Spotlighting, StruQ, SecAlign via LoRA adapter) and 5 are reimplemented from the papers' descriptions (Task Shield, Attention Tracker, IPIGuard, FATH, Polymorphic Prompt, Mixture-of-Encodings). The reimplementations follow the algorithms described in the original papers but may differ in implementation details (e.g., exact prompt templates, threshold values). We note this as a limitation: our ASR numbers for reimplemented defenses reflect our implementations, not the original authors' code. The framework is designed to accept original implementations via the registry interface, and we encourage defense authors to integrate their code for more faithful evaluation.

\paragraph{Reproducibility.}
All code, configurations, and results are available at the project repository (anonymized for double-blind review). The framework runs end-to-end with a stub backend for CI and with OpenAI-compatible APIs for real evaluation. Checkpointing supports resuming interrupted runs.
