\section{Related Work}
\label{sec:related}

\paragraph{Indirect Prompt Injection.}
Greshake et al.~\cite{greshake2023not} first demonstrated that LLMs cannot distinguish instructions from data when both appear in the context window, enabling IPI. InjecAgent~\cite{zhan2024injecagent} systematized the attack with 1{,}054 injection cases across 17 scenarios. AgentDojo~\cite{debenedetti2024agentdojo,debenedetti2025agentdojo} provided a dynamic evaluation environment with 973 tasks and built-in injection mechanisms. WebInject~\cite{wang2025webinject} extended IPI to multimodal web agents. AgentVigil~\cite{wang2025agentvigil} introduced black-box red-teaming for IPI. MUZZLE~\cite{syros2026muzzle} proposed adaptive agentic red-teaming against web agents. Our framework unifies these attack methodologies into a composable generator with per-defense evasion strategies.

\paragraph{Defenses Against IPI.}
Spotlighting~\cite{hines2024defending} introduced input encoding; StruQ~\cite{chen2024struq} and SecAlign~\cite{chen2024secalign} proposed structured queries and preference optimization; Task Shield~\cite{jia2025task} enforced task alignment; Attention Tracker~\cite{hung2025attention} used internal probing; IPIGuard~\cite{an2025ipiguard} built tool-dependency graphs; FATH~\cite{wang2024fath} used authentication tokens; MELON~\cite{zhu2025melon} offered provable defense; IsolateGPT~\cite{wu2025isolategpt} and ACE~\cite{li2025security} proposed architecture-level isolation; OpenClaw~\cite{cheng2026agent} proposed agent privilege separation. Our taxonomy organizes these into three evaluated classes (IE, SQ, RC) and evaluates them end-to-end under both static and adaptive attack.

\paragraph{Adaptive Attacks.}
Zhan et al.~\cite{zhan2025adaptive} showed that adaptive attacks break Spotlighting, StruQ, and SecAlign. Our work extends this to 11 defenses evaluated end-to-end with native AgentDojo evaluators, comparing static (L1) vs.\ adaptive (L2) attacks under real tool execution. Chen et al.~\cite{chen2025backdoorpowered} showed that backdoor-augmented injection nullifies defense methods.

\paragraph{Concurrent Surveys.}
Gulyamov et al.~\cite{gulyamov2026prompt} provide a comprehensive review of prompt injection vulnerabilities and defense mechanisms. Geng et al.~\cite{geng2025prompt} survey attack methods, root causes, and defense strategies. Unlike these surveys, our work is an empirical SoK with a native adaptive attack framework and end-to-end evaluation using AgentDojo's native security and utility evaluators; the surveys are taxonomic and do not conduct adaptive-attack evaluations.

\paragraph{LLM Agent Security.}
Beyond IPI, recent work studies tool-selection attacks~\cite{shi2025prompt}, cross-tool pollution, over-privileged tool selection, multi-agent injection propagation, and LLM agent system architectures. Our SoK focuses specifically on the IPI attack-defense ecosystem and complements these broader surveys.
