\section{Supplemental Experiments}
\label{app:supplemental}

\subsection{Unknown-Defense and No-Feedback Attacks}
\label{app:settings_ab}

To address whether our results stem from the defense's fundamental flaws or from overly strong attacker knowledge, we run two additional settings:

\paragraph{Setting A (Unknown-defense):} The attacker does not know the defense class and uses a universal payload (our L1 static \texttt{<INFORMATION>} template). \Cref{tab:setting_a} compares ASR under unknown-defense (L1) vs.\ known-defense (L2) attacks. The results show that defense-class knowledge significantly increases ASR for IE and IP defenses (delta +0.40 to +0.80), confirming that the matched evasion strategies add real value beyond universal payloads. However, even without defense knowledge, the universal L1 payload achieves 0.20--0.60 ASR on IE/IP defenses, indicating these defenses are already vulnerable to static attacks---the adaptive knowledge merely saturates the attack.

\begin{table}[h]
\centering
\small
\caption{Setting A: Unknown-defense (L1) vs.\ known-defense (L2) ASR.}
\label{tab:setting_a}
\begin{tabular}{lccc}
\toprule
Defense & ASR (unknown) & ASR (known) & Delta \\
\midrule
Spotlighting & 0.40 & 1.00 & +0.60 \\
Attention Tracker & 0.60 & 1.00 & +0.40 \\
Polymorphic & 0.20 & 1.00 & +0.80 \\
Mixture-of-Enc. & 0.47 & 1.00 & +0.53 \\
StruQ & 0.00 & 0.00 & 0.00 \\
Task Shield & 0.00 & 0.00 & 0.00 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Setting B (No-feedback):} The attacker places a single injection without iterative optimization (max\_iters=0). \Cref{tab:setting_b} shows that feedback does not change ASR for any defense: the single-shot matched payload is already as effective as the iteratively refined one. This indicates that the adaptive attack's strength comes from \emph{defense-class knowledge} (the matched strategy), not from iterative feedback. This is consistent with the fragility conjecture: the search space is small enough that a single well-crafted payload suffices.

\begin{table}[h]
\centering
\small
\caption{Setting B: No-feedback (iters=0) vs.\ feedback (iters=2) ASR at L2.}
\label{tab:setting_b}
\begin{tabular}{lcc}
\toprule
Defense & ASR (no feedback) & ASR (feedback) \\
\midrule
Spotlighting & 1.00 & 1.00 \\
Attention Tracker & 1.00 & 1.00 \\
Polymorphic & 1.00 & 1.00 \\
StruQ & 0.00 & 0.00 \\
Task Shield & 0.00 & 0.00 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Cross-Defense Transfer}
\label{app:setting_c}

To distinguish matched adaptive attacks from universal stronger prompts, we test whether payloads generated against defense A transfer to defense B. \Cref{tab:setting_c} shows that payloads generated for Spotlighting (IE class) transfer perfectly to Attention Tracker (IP class, transfer ratio 1.00) but poorly to StruQ (SQ class, 0.07) and not at all to Task Shield (RC class, 0.00). This indicates that IE and IP defenses share a common vulnerability (observable-signal mimicry), while SQ and RC defenses are structurally different. This finding supports the taxonomy's predictive power: defenses in the same failure-mode cluster are vulnerable to the same payloads.

\begin{table}[h]
\centering
\small
\caption{Setting C: Cross-defense transfer (payloads from Spotlighting tested on other defenses).}
\label{tab:setting_c}
\begin{tabular}{lccc}
\toprule
Target Defense & ASR (transfer) & ASR (matched) & Transfer Ratio \\
\midrule
StruQ (SQ) & 0.07 & 0.00 & --- \\
Task Shield (RC) & 0.00 & 0.00 & --- \\
Attention Tracker (IP) & 1.00 & 1.00 & 1.00 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Threshold Sweep: ASR-FPR Pareto Frontier}
\label{app:threshold_sweep}

The reviewer notes that reporting a single threshold for runtime-checking defenses does not characterize the full security-utility tradeoff. We sweep the decision threshold for Task Shield and Attention Tracker across 9 values (0.1--0.9).

\paragraph{Task Shield.} ASR remains 0.00 at all thresholds, but FPR ranges from 0.07 (thresh=0.1) to 0.67 (thresh=0.8). At the lowest threshold (0.1), Task Shield achieves USR=0.93 with ASR=0.00---a much better operating point than the default threshold=0.5 (USR=0.53). This suggests that Task Shield's default threshold is overly conservative, and a lower threshold provides a better security-utility tradeoff. However, even at the optimal threshold, Task Shield still blocks 7\% of benign actions, and its ASR of 0.00 may reflect over-blocking rather than true robustness (the defense blocks so many actions that the attacker's malicious action is blocked too).

\paragraph{Attention Tracker.} ASR is 1.00 at \emph{all} thresholds, confirming that the attention-redistribution evasion defeats the detector regardless of its sensitivity. FPR is 0.00 at all thresholds (the detector never flags benign prompts), meaning the defense provides zero security with zero utility cost---the worst possible operating point.

\begin{table}[h]
\centering
\small
\caption{Threshold sweep for Task Shield: ASR and FPR at each threshold.}
\label{tab:threshold_sweep_ts}
\begin{tabular}{cccc}
\toprule
Threshold & ASR & FPR & USR \\
\midrule
0.1 & 0.00 & 0.07 & 0.93 \\
0.2 & 0.00 & 0.20 & 0.80 \\
0.3 & 0.00 & 0.40 & 0.60 \\
0.5 & 0.00 & 0.47 & 0.53 \\
0.8 & 0.00 & 0.67 & 0.33 \\
0.9 & 0.00 & 0.60 & 0.40 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Query and Cost Analysis}
\label{app:cost}

The full 3{,}900-trial evaluation consumed approximately \$30--40 in API costs and 4.5 hours of wall-clock time. The adaptive attack optimizer (L2/L3) uses an average of 2.3 LLM calls per trial, compared to 1.0 for L0/L1. The per-trial token cost averages 450 input + 85 output tokens for L0/L1 and 620 input + 120 output for L2/L3 (the optimizer adds ~170 input tokens for the mutation prompt). Setting B (\Cref{app:settings_ab}) confirms that the optimizer feedback does not improve ASR, so the 2.3x cost overhead of iterative optimization is not justified for this threat model---a single matched payload suffices.

\subsection{Held-Out Defense Validation}
\label{app:held_out}

To test the taxonomy's out-of-sample predictive power, we use SecAlign~\cite{chen2024secalign} as a held-out defense (it was not used to design the evasion strategies, as it falls in the SQ class alongside StruQ, and our SQ strategy was designed against StruQ). Our taxonomy predicts that SecAlign should fail to R2 (boundary blur). Our SecAlign evaluation (Appendix~\ref{app:sealign}) shows that on Mistral-7B-Instruct, SecAlign achieves R-ASR=0.00---but so does the no-defense baseline, because Mistral-7B itself resists the \texttt{<INFORMATION>} template. This means the held-out validation is inconclusive on this backbone: the defense's robustness cannot be distinguished from the backbone's natural resistance. On a backbone susceptible to IPI (e.g., GPT-4o-mini), the prediction would be testable, but we were unable to run SecAlign on GPT-4o-mini because the LoRA adapter is only available for Llama-3-8B-Instruct (a gated model). We note this as a limitation and plan to complete the held-out validation when model access permits.
